Empirical evidence for positive selection that is not adaptive evolution

### 2022-04-22 manuscript/main.tex
Manuscript by JJ and TL.

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\BiDimArray}[1]{\bm{#1}}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathrm{Var}}
\newcommand{\der}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\avg}[1]{\left< #1 \right>} % for average
\newcommand{\Ne}{N_{\mathrm{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\mathrm{fix}}}
\newcommand{\Spop}{\beta}
\newcommand{\SpopMean}{\overline{\Spop}}
\newcommand{\Sphy}{S}
\newcommand{\SphyMean}{\overline{\Sphy}}
\newcommand{\polyDel}{\proba \left[ {\Spop < -1} \right]}
\newcommand{\polyNeutral}{\proba \left[ -1 < \Spop < 1 \right]}
\newcommand{\polyAdv}{\proba \left[ 1 < \Spop  \right]}

@@

    \begin{abstract}
        Evidence for positive selection occurring in protein-coding DNA sequence is widespread across different taxa.
        Positive selection can be a result of adaptive evolution, meaning that individuals have had to adapt to a change in environment or selective pressure.
        However, current positive selection can also be the result of mutations compensating for deleterious substitutions that have accumulated along lineages, in which case positive selection is non-adaptive and predictable.
        To evaluate the extent of non-adaptive positive selection, we have combined population-genetics and phylogenetics datasets.
        We first leveraged phylogenetic codon models which are based on a population-genetics formalism, assuming a non-adaptive fitness landscape.
        These models estimate the fitness of each of the 20 amino acids for each protein site, given mammalian protein-coding DNA alignments and gene tree topologies.
        Second, we integrated mammalian divergence data with polymorphic variants found in 29 populations across 7 mammalian genera.
        For each non-synonymous variant observed at the population level, we predicted its change in fitness from amino-acid fitnesses estimated at the mammalian scale.
        We found that a large proportion of observed non-synonymous changes are predicted to be positively selected, meaning that the ancestral allele is sub-optimal.
        These supposedly advantageous variants are indeed showing signs of recent positive selection in all populations.
        Our work confirms that deleterious substitutions have accumulated across the phylogeny and are currently being compensated for, resulting in widespread positive selection that is not adaptive evolution.
        This study also shows the leverage obtained by integrating phylogenetic and population genetics on a common formalism.
    \end{abstract}

@@

    We usually say that a mutation is undergoing positive selection when it brings a fitness advantage to its bearer, and thus facilitate its own transmission to the next generation.
    However, the causes of this advantage can be diverse.
    We can distinguish two main advantages.
    The first correspond to what we usually call adaptive evolution.
    he advantage in fitness of a new mutation allows its bearer to become better adapted to a changing environment.
    The second we call maintenance positive selection.
    The mutation restores the ancestral state which corresponded to an optimal functioning of the individual in a constant environment, but has been lost due to the fixation of a deleterious mutation by drift.
    These two different kinds of selection result in a positive transmission bias of the advantageous allele, but their consequences are conceptually different.
    While adaptive evolution promotes phenotype diversification, maintenance positive selection promotes phenotype stability and preserves well established biological systems.
    It is thus of importance to be able to distinguish those two kinds of positive selection in order to understand the diversification of species.

    Adaptive evolution in proteins can be quantified by comparing sequence variation inside a population to the sequence divergence with a sister species\cite{mcdonald_adaptative_1991}.
    Indeed, strongly adaptive mutations are expected to be observed in substitutions but not in polymorphism because they reach fixation quickly.
    This method has been largely improved to account for possible bias-inducing factors such as the presence of weak positive selection, background selection and demography fluctuations\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016, tataru_inference_2017}.
    It's application across many taxa has uncovered that a large proportion of substitutions are positively selected\cite{moutinho_variation_2019}.
    The root causes for this widespread rate of positive selection are however scarcely investigated, often assumed to be the end result of adaptive evolution.
    However, as mentioned above, not all these mutations contribute to genetic innovation and diversification.
    Adaptive evolution can only be well estimated if maintenance positive selection is taken into account.
    While adaptive evolution is unpredictable because it brings an advantage concerning an unpredictable change in the environment of organisms, the second is predictable because optimal states should be more or less the same for closely related organisms which share most of their environment and physiology.
    A way to measure maintenance positive selection is to compute selection coefficients assuming a constant and stable fitness landscape estimated from the alignment of protein-coding genes.
    For each site of the proteic alignment, the fitness of each amino-acid can be estimated from their frequency in the alignment.
    One of the most used method to make these predictions is SIFT (Sort Intolerant From Tolerant) which reliably detects deleterious mutations in many taxa.
    However, this method has two major shortcomings.
    First, this method uses amino-acid frequency as a proxy for amino-acid fitness, even if we know that a lot of mechanisms can influence amino-acid frequency other than fitness (mutation biases, GC-biased gene conversion, mutational paths).
    Second, each mutations is attributed a score to describe if it should be under purifying selection on one hand or be neutral or advantageous on the other hand.
    But this score makes little sense to actual measure of fitness, and is not interpretable in terms of population genetics.
    So called mutation-selection codon models provide a nearly-neutral framework of evolution by estimating the fitness landscape over amino-acid sequences for each site of the sequence while taking into account non adaptive-processes impacting codon frequency\cite{ halpern_evolutionary_1998, rodrigue_mechanistic_2010, tamuri_estimating_2012}.
    Only nearly-neutral mutations between high fitness amino acids will tend to be permitted by the model, allowing for the explicit calculation of the scaled selection coefficient of non-synonymous mutations between codons.
    These models have been used as a null-model to predict the overall rate of evolution of proteins\cite{spielman_relationship_2015, dosreis_how_2015}.
    A departure from this model indicates that proteins are evolving under a changing fitness landscape\cite{rodrigue_detecting_2017, tamuri_mutationselection_2021} and is a signal of adaptive evolution\cite{rodrigue_bayesian_2021}.
    In this study we want to address three main questions: which proportions of the polymorphism we observe in mammals is undergoing maintenance positive selection ?
    Are these mutations actually under positive selection at the scale of populations ?
    How does effective population size modulate positive, negative and weak selection ?

@@

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure1}
        \caption{
            Integrating divergence and polymorphism for the detection of adaptation.
            At the phylogenetic level (panel A), amino-acid Wrightian fitness for each site are estimated from protein-coding DNA alignments using mutation-selection codon models.
            At the population-genetic level (panel B), for each observed single nucleotide polymorphism (SNP), selection coefficient are computed as the difference in amino-acid fitness between the ancestral and derived variant.
        }
        \label{fig:method}
    \end{figure*}

    % How to articulate phylogenetic and population genetic
    We developed and leveraged a pipeline integrating divergence and polymorphism data across the entire exome for 29 populations across 7 genera, namely \textit{Equus}, \textit{Canis}, \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus} and \textit{Homo} (fig.~\ref{fig:method}).
    For each non-synonymous SNP observed at the population level, we predicted its scaled selection coefficient ($\Sphy$) in fitness from amino-acid fitness estimated at the mammalian scale (fig.~\ref{fig:homo-afr-results}A).
    Then we ask whether the variants that are supposedly advantageous are showing signature of short term positive selection.
    In other words, if a variant is found in many mammals but not in ancestral humans, and a mutation in human is toward the mammal version, is this mutation advantageous for the individual carrying it?
    We divided all observed SNPs in 5 classes of selection: severely deleterious with $\Sphy<-3$, deleterious with $-3<\Sphy<-1$, weakly deleterious with $-1<\Sphy<0$, weakly advantageous with $0<\Sphy<1<$ and advantageous with $1<\Sphy$.
    For each class of selection, we first derived its site-frequency spectrum (SFS), which represent the proportion of mutations with a given number of derived allele in the population (fig.~\ref{fig:homo-afr-results}B).
    Compared to synonymous mutations (supposedly neutral), mutations predicted to be advantageous are over-represented at higher frequency, while severally deleterious are under-represented at high frequency, a signature that our prediction are qualitatively in the right direction.
    Aside from the frequency of the derived allele, another information is the total number of observed alleles compared to its expectation, given by the total mutation rate.
    The total mutation rate for each class is computed from the ancestral genome reconstruction (fig.~\ref{fig:homo-afr-results}C).
    For example, for the category of selection with $1<\Sphy$, the expected total mutation rate is obtained by adding together mutation rates for all possible non-synonymous mutations for which the derived allele is toward a fitter amino-acid than the ancestral allele with $\Sphy > 1$.
    The total mutation rate and the SFS are combined to estimate the distribution of the selection coefficient $\Spop$ given by population-based model\cite{tataru_inference_2017, tataru_polydfe_2020}.
    These population based models cannot give a specific estimate of $\Spop$ for each SNP, but by gathering information across many SNPs, these models can estimate the proportion of advantageous mutations ($\polyAdv$), of nearly-neutral mutations ($\polyNeutral$) and of deleterious mutations ($\polyDel$) (fig.~\ref{fig:homo-afr-results}D).
    We found that SNPs predicted to be advantageous given their selection coefficient $\Sphy$ are indeed under positive selection at the population-genetic scale, confirming that there is a local ongoing selection toward amino acids that are advantageous across mammals.
    From another angle, mutations toward amino acids deemed deleterious across mammals are effectively purged at the population-genetic scale.
    In between, mutations predicted to be weakly selected are effectively composed of neutral, positively and negatively selected mutations.

    We replicated such experiment across 29 populations (fig.~\ref{fig:all-pop-results}), and found that mutations predicted to be deleterious are indeed purged at the population-genetic scale (fig.~\ref{fig:all-pop-results}A), while mutations predicted to be advantageous are indeed positively selected (fig.~\ref{fig:all-pop-results}C) .
    However, mutation predicted to be weakly advantageous ($0<\Sphy<1<$) are composed of a mix of neutral and selected mutations with varying proportions across the different populations (fig.~\ref{fig:all-pop-results}B).
    When populations are ordered by increased diversity (fig.~\ref{fig:selcoeff-pop-phylo}A), the proportion of neutral mutations is decreasing, suggesting that populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}) are more likely to discriminate whether mutations are advantageous or deleterious.
    Alternatively stated, mutations in populations with low diversity (e.g.~\textit{Homo}) are effectively nearly-neutral and behave as would a neutral mutation.
    This result is in accordance with the nearly-neutral theory of evolution that mutations are less efficiently selected for small populations.
    Moreover, we found this effect to be in the same direction but less pronounced when we focused on all SNPs (fig.~S1) than on the SNPs predicted to be weakly selected (fig.~S2-3).
    For populations with high diversity, because mutations are less likely to be neutral they are mechanically more likely to be selected and thus deleterious (fig.~S4), especially for mutations predicted to be deleterious at the phylogenetic scale (fig.~S5-6).
    However, and surprisingly, a decrease in the proportion of neutral mutation did not translate into an increase proportion of advantageous mutations.
    Instead, populations with higher diversity have both a lower proportion of neutral and advantageous mutations (fig.~S7), this effect is also more pronounced for mutations predicted to be advantageous at the phylogenetic scale (fig.~S8).
    This discrepancy can be explained by the fact that population with larger diversity have already depleted the putatively advantageous mutations, and that only deleterious are remaining possible.
    In other words, being on the top of a fixed fitness landscape leaves you with only paths going downhill.
    Altogether, populations with increased diversity are showing signature of increased selection efficiency and are also less likely to generate positively selected mutations.
    To look further into the relation between diversity and selection efficacy, we studied the estimation of the mean $\SpopMean$ as a function of the mean $\SphyMean$ for a subset of SNPs, for which $\SphyMean$ is close to $0$.
    Sliding window of groups of $5.000$ SNPs are taken with a mean $\SphyMean$ between $-0.5$ and $0.5$ (fig.~\ref{fig:selcoeff-pop-phylo}B).
    $\SpopMean$ is estimated for each group, with some groups that might have part of same SNPs that overlap.
    The relationship between $\SpopMean$ and $\SphyMean$ is described as a second order polynomial, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\SpopMean/\SphyMean$ for each population.
    We observe that $\SpopMean/\SphyMean$, representing the efficiency of selection at the short versus long term is positively correlated with synonymous diversity (fig.~\ref{fig:selcoeff-pop-phylo}C), a proxy of effective population size.
    Altogether, population with large effective population size are under more efficient selection, confirming the prediction of the nearly-neutral theory of evolution.

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure2}
        \caption{
            Panel A: Histogram of predicted selection coefficient (S) for each observed mutation in a sample of 8 individuals (out of 512 in the original dataset) with African descent.
            Mutations are divided in 5 classes of selection: severely deleterious (blue), deleterious (green), weakly deleterious (light green), weakly advantageous (yellow) and advantageous (red).
            Panel B: Site-frequency spectrum (SFS) represent the proportion of mutations (y-axis) with a given number of derived allele in the population (x-axis).
            SFS are drawn for a random sample of 16 alleles (mean in solid line and standard deviation in filled color) for each class of selection and for synonymous mutations supposedly neutral (black).
            Supposedly advantageous are over-represented at higher frequency, while severally deleterious are under-represented at high frequency, a signature of selection.
            Panel C. Histogram of predicted selection coefficient (S) for each possible mutation away from the ancestral human genome, weighted by the mutation rate.
            For each class of selection, the sum of histogram in this class thus gives the expected total mutation rate toward this class.
            If they are less mutations observed than expected, this class is thus undergoing purifying selection.
            Panel D. For each class of selection (and for the set of all non-synonymous mutations), information from the SFS and the expected total mutation rate are combined at the population-genetic scale to estimate the proportion of advantageous mutations $\polyAdv$, of nearly-neutral mutations $\polyNeutral$ and of deleterious mutations $\polyDel$.
        }
        \label{fig:homo-afr-results}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure3}
        \caption{
            Reproducibility of results shown in figure~\ref{fig:homo-afr-results} in 29 populations across 7 genera.
            Site-frequency spectra and total mutation rates are combined at the population-genetic scale to estimate the proportion of advantageous mutations ($\polyAdv$) in top panel, of nearly-neutral mutations ($\polyNeutral$) in middle panel and of deleterious mutations ($\polyDel$) in bottom panel.
        }
        \label{fig:all-pop-results}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure4}
        \caption{
            Panel A. For each population, proportion of advantageous ($\polyAdv$), nearly-neutral ($\polyNeutral$) and of deleterious ($\polyDel$) mutations estimated at the population-genetic scale.
            Estimation is performed with SNPs predicted to be weakly-advantageous at the phylogenetic scale ($0<\Sphy<1$).
            Panel B. Mean selection coefficient at the population scale $\SpopMean$ as a function of phylogenetic $\SphyMean$.
            Observed SNPs are sorted and groups of $5.000$ SNPs are taken as a sliding windows with increasing mean $-0.5 \leq \SphyMean \leq 0.5$.
            For each group along the sliding window, $\SpopMean$ (y-axis) is estimated based on SFS and the total mutation rate, and compared to the mean $\SphyMean$ (x-axis).
            A second order polynomial is fitted for each population, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\SpopMean/\SphyMean$ given in legend.
            Panel C. Ratio $\SpopMean/\SphyMean$ as a function of the synonymous diversity for each population.
            Populations with higher diversity are showing higher ratio of ratio $\SpopMean/\SphyMean$.
        }
        \label{fig:selcoeff-pop-phylo}
    \end{figure*}

    % Our study is a bridge between phylogenetic and population genetics
    From a theoretical evolution perspective, we found that we could use phylogenetic to predict the selection coefficient of a mutation.
    Appart from phylogenetic mutation-selection codon models, SIFT scores based on amino-acid alignments across species have already be found to be informative of the selection pressure exerted at the population scale\cite{chen_hunting_2021}.
    However, they do not integrate the underlying phylogeny nor the underlying mutational process.
    Additionally, they are not based on population-genetics formalism, hence the relationship between selection coefficient and SIFT score is not straightforward.
    We found here that mutation-selection formalism greatly outperforms SIFT score to predict the selection coefficient of mutations (Supplementary Materials).
    Most importantly, we showed that mutation-selection codon models formalize a bridge between phylogenetics and population genetics.
    Our study is a bridge between phylogenetic and population genetics, but both ends has its limitation.

>>>>

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\BiDimArray}[1]{\bm{#1}}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathrm{Var}}
\newcommand{\der}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\avg}[1]{\left< #1 \right>} % for average
\newcommand{\Ne}{N_{\mathrm{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\mathrm{fix}}}

\newcommand{\Sphy}{S}
\newcommand{\SphyMean}{\overline{\Sphy}}
\newcommand{\divStrongDel}{\Sphy < -3}
\newcommand{\divDel}{-3 < \Sphy < -1}
\newcommand{\divWeakDel}{-1 < \Sphy < 0}
\newcommand{\divWeakAdv}{0 < \Sphy < 1}
\newcommand{\divAdv}{ \Sphy > 1}
\newcommand{\PdivStrongDel}{\proba \left[ \divStrongDel \right]}
\newcommand{\PdivDel}{\proba \left[ \divDel \right]}
\newcommand{\PdivWeakDel}{\proba \left[ \divWeakDel \right]}
\newcommand{\PdivWeakAdv}{\proba \left[ \divWeakAdv \right]}
\newcommand{\PdivAdv}{\proba \left[ \divAdv \right]}

\newcommand{\Spop}{\beta}
\newcommand{\SpopMean}{\overline{\Spop}}
\newcommand{\polyDel}{\Spop < -1}
\newcommand{\polyNeutral}{-1 < \Spop < 1}
\newcommand{\polyAdv}{ \Spop > 1}
\newcommand{\PpolyDel}{\proba \left[ \polyDel \right]}
\newcommand{\PpolyNeutral}{\proba \left[ \polyNeutral \right]}
\newcommand{\PpolyAdv}{\proba \left[ \polyAdv \right]}

@@

    \begin{abstract}
        Evidence for positive selection occurring in protein-coding DNA sequence is widespread across different taxa.
        Positive selection can be a result of adaptive evolution, meaning that individuals have had to adapt to a change in environment or selective pressure.
        However, current positive selection can also be the result of mutations compensating for deleterious substitutions that have accumulated along lineages, in which case positive selection is non-adaptive and predictable.
        To evaluate the extent of non-adaptive positive selection, we have combined population genetics and phylogenetic datasets.
        We first leveraged phylogenetic codon models which are based on a population-genetics formalism, assuming a non-adaptive fitness landscape.
        These models estimate the fitness of each of the 20 amino acids for each protein site, given mammalian protein-coding DNA alignments and gene tree topologies.
        Second, we integrated mammalian divergence data with polymorphic variants found in 29 populations across 7 mammalian genera.
        For each non-synonymous variant observed at the population level, we predicted its change in fitness from amino-acid fitness estimated at the mammalian scale.
        We found that a large proportion of observed non-synonymous changes are predicted to be positively selected, meaning that the ancestral allele is suboptimal.
        These supposedly advantageous variants are indeed showing signs of recent positive selection in all populations.
        Our work confirms that deleterious substitutions have accumulated across the phylogeny and are currently being compensated for, resulting in widespread positive selection that is not adaptive evolution.
        This study also shows the leverage obtained by integrating phylogenetic and population genetics on a common formalism.
    \end{abstract}

@@

    We usually say that a mutation is undergoing positive selection when it brings a fitness advantage to its bearer, and thus facilitate its own transmission to the next generation.
    However, the causes of this advantage can be diverse.
    We can distinguish two main advantages.
    The first advantage allows its bearer to become better adapted to a new or a changing environment, what we typically call adaptive evolution.
    The population as a whole is thus climbing the fitness landscape along an adaptive walk\cite{tenaillon_utility_2014}.
    The second is non-adaptive, in the sense that the increase in fitness of a new mutation is not a response to a new or a changing environment.
    Instead, the mutation restores the ancestral state which corresponded to an optimal functioning of the individual in a constant environment, but has been lost due to the fixation of a deleterious mutation by drift.
    The population as a whole is thus merely staying in the same position in fitness landscape at this equilibrium, and positive selection compensates for deleterious mutations that reach fixation due to drift\cite{sella_application_2005, mustonen_fitness_2009}.
    These two different kinds of selection result in a positive transmission bias of the advantageous allele, but their consequences are conceptually different.
    While adaptive evolution promotes phenotype diversification, non-adaptive positive selection promotes phenotype stability and preserves well established biological systems.
    It is thus of importance to be able to distinguish those two kinds of positive selection in order to understand the diversification of species.

    Inside proteins, positive selection can be quantified by comparing sequence variation inside a population with the sequence divergence with a sister species\cite{mcdonald_adaptative_1991}.
    Indeed, strongly advantageous mutations are expected to be observed in substitutions but not in polymorphism because they reach fixation quickly.
    This method has been largely improved to account for possible bias-inducing factors such as the presence of weak positive selection, background selection and demography fluctuations\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016, tataru_inference_2017}.
    It's application across many taxa has uncovered that a large proportion of substitutions are positively selected\cite{moutinho_variation_2019}.
    The root cause for this widespread rate of positive selection is, however, scarcely investigated, often assumed to be the end result of adaptive evolution.
    However, as mentioned above, not all these mutations contribute to genetic innovation and diversification.
    Adaptive evolution can only be well estimated if non-adaptive positive selection is taken into account.
    While adaptive evolution is unpredictable because it brings an advantage concerning an unpredictable change in the environment of organisms, the second is predictable because optimal states should be more or less the same for closely related organisms which share most of their environment and physiology.
    A way to thus measure non-adaptive positive selection is to compute selection coefficients assuming a constant and stable fitness landscape estimated from the alignment of protein-coding genes.
    For each site of the protein alignment, amino-acid frequencies in the alignment is a proxy for its fitness, since for example a deleterious amino acid will be absent in the alignment.
    One of the most used methods to make these predictions is SIFT (Sort Intolerant From Tolerant) which reliably detects deleterious mutations in many taxa\cite{ng_sift_2003, vaser_sift_2016}.
    However, this method has two major shortcomings.
    First, this method uses amino-acid frequency as a proxy for amino-acid fitness, even if we know that a lot of mechanisms can influence amino-acid frequency other than fitness (phylogeny, mutation biases, GC-biased gene conversion, mutational paths).
    Second, each mutation is attributed a score to describe if it should be under purifying selection on one hand or be neutral or advantageous, on the other hand.
    But this score makes little sense to actual measure of fitness, and is not interpretable in terms of population genetics.
    In the other hand, at the DNA level, so-called mutation-selection codon models are rooted with population genetics formalism\cite{halpern_evolutionary_1998, mccandlish_modeling_2014}.
    Such models provide a nearly-neutral framework of evolution by estimating the fitness landscape over amino-acid sequences for each site of the sequence while taking into account non-adaptive processes impacting codon frequency\cite{halpern_evolutionary_1998, rodrigue_mechanistic_2010, tamuri_estimating_2012}.
    Only nearly-neutral mutations between high fitness amino acids will tend to be permitted by the model, allowing for the explicit calculation of the scaled selection coefficient of non-synonymous mutations between codons.
    These models have been used as a null model to predict the overall rate of evolution of proteins\cite{spielman_relationship_2015, dosreis_how_2015}.
    A departure from this model indicates that proteins are evolving under a changing fitness landscape\cite{rodrigue_detecting_2017, tamuri_mutationselection_2021} and is a signal of adaptive evolution\cite{rodrigue_bayesian_2021}.
    In this study we want to address three main questions: which proportions of the polymorphism we observe in mammals is undergoing non-adaptive positive selection ?
    Are these mutations actually under positive selection at the scale of populations ?
    How does effective population size modulates positive, negative and weak selection ?

    \begin{figure*}[!ht]
        \
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure1}
        \caption{
            Integrating divergence and polymorphism for the detection of adaptation.
            At the phylogenetic level (panel A), amino-acid Wrightian fitness for each site are estimated from protein-coding DNA alignments using mutation-selection codon models.
            At the population-genetic level (panel B), for each observed single nucleotide polymorphism (SNP), selection coefficient are computed as the difference in amino-acid fitness between the ancestral and derived variant.
        }
        \label{fig:method}
    \end{figure*}

@@

    We developed and leveraged a pipeline integrating divergence and polymorphism data across the entire exome for 29 populations across 7 genera, namely \textit{Equus}, \textit{Canis}, \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus} and \textit{Homo} (fig.~\ref{fig:method}).
    For each non-synonymous SNP observed at the population level, we predicted its scaled selection coefficient ($\Sphy$) in fitness from amino-acid fitness estimated at the mammalian scale (fig.~\ref{fig:homo-afr-results}A).
    Then we ask whether the variants that are supposedly advantageous are showing signature of short term positive selection.
    In other words, if a variant is found in many mammals but not in ancestral humans, and a mutation in humans is toward the mammalian version, is this mutation advantageous for the individual carrying it?
    At the population level (e.g. in humans with Asian descent in fig.~\ref{fig:homo-afr-results}), we divided all observed SNPs in 5 classes of predicted selection coefficient: severely deleterious with $\divStrongDel$, deleterious with $\divDel$, weakly deleterious with $\divWeakDel$, weakly advantageous with $\divWeakAdv$ and advantageous with $\divAdv$.
    For each class of selection, SNPs are then grouped by the number of derived alleles in the population (fig.~\ref{fig:homo-afr-results}B), generating so called site-frequency spectrum (SFS).
    Compared to synonymous mutations supposedly neutral (black line in fig.~\ref{fig:homo-afr-results}B), mutations predicted to be deleterious (blue and green lines in fig.~\ref{fig:homo-afr-results}B) are underrepresented at high frequency in the population, a signature of ongoing purifying selection.
    Remarkably, SNPs predicted to be advantageous (red line in fig.~\ref{fig:homo-afr-results}B) are over-represented at higher frequency in the population, meaning they are more likely to reach fixation as a consequence of ongoing positive selection.
    However, these observations remain qualitative.
    Thus, we quantified the distribution of selection coefficient for each category of SNPs with the model polyDFE\cite{tataru_inference_2017, tataru_polydfe_2020}.
    The model uses the count of derived allele for each frequency class of the SFS to infer the distribution of fitness effects (DFE) of the region under study.
    For each category of SNPs, it is important to compute the expected number of mutations giving birth to a SNP of that category given a mutational matrix (fig.~\ref{fig:homo-afr-results}C).
    The model uses both the mutational expectation for each category, and the counts of synonymous derived allele as a neutral model, which allow to control for demographic effects, and polarization errors when inferring the ancestral state.
    These populations based models cannot give a specific estimate of $\Spop$ for each SNP, but by gathering information across many SNPs, these models can estimate the proportion of advantageous mutations ($\PpolyAdv$), of nearly-neutral mutations ($\PpolyNeutral$) and of deleterious mutations ($\PpolyDel$) (fig.~\ref{fig:homo-afr-results}D).
    We found that SNPs predicted to be deleterious given their selection coefficient $\Sphy$ across mammals are effectively purged at the population-genetic scale, demonstrating that purifying selection is largely predictable and amino acids with low fitness across mammals are effectively purged away in each population.
    Notably, SNPs predicted to be advantageous given their selection coefficient $\Sphy$ are indeed under positive selection at the population-genetic scale, confirming that there is a local ongoing selection toward amino acids that are restoring the optimally functioning protein.
    In between these two extremes, mutations predicted to be weakly selected (cyan and orange in fig.~\ref{fig:homo-afr-results}) are effectively composed of neutral, positively and negatively selected mutations.

    We replicated such experiment across 29 populations (fig.~\ref{fig:all-pop-results}), and found that mutations predicted to be deleterious are indeed purged at the population-genetic scale (blue and green rows in fig.~\ref{fig:all-pop-results}A).
    However, mutation predicted to be weakly advantageous ($\divWeakAdv$) are composed of a mix of neutral and selected mutations with varying proportions across the different populations (orange row in fig.~\ref{fig:all-pop-results}B).
    Similarly, mutations predicted to be advantageous are found to be positively selected with varying proportions (red row in fig.~\ref{fig:all-pop-results}C).
    A variable which could explain those variations between populations is the effective number of individuals in the population ($\Ne$), a major driver of selection efficacy.
    $\Ne$ is not directly accessible, and a reasonable proxy usually used is the synonymous diversity ($\theta=4\Ne\mu$ with $\mu$ the mutation rate).
    We thus ordered populations by increased synonymous diversity (fig.~\ref{fig:selcoeff-pop-phylo}A), exhibiting a decreased proportion of neutral mutations.
    This result suggest that populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}) are more likely to discriminate whether mutations are advantageous or deleterious.
    Alternatively stated, mutations in populations with low diversity (e.g.~\textit{Homo}) are effectively nearly-neutral and behave as would a neutral mutation.
    This result is qualitatively in accordance with the nearly-neutral theory of evolution that mutations are less efficiently selected for small populations.
    Moreover, we found this effect to be in the same direction but less pronounced when we focused on all SNPs (fig.~S1) rather than only the SNPs predicted to be weakly selected (fig.~S2-3).
    To look further into the relation between diversity and selection efficacy, we thus studied the estimation of the mean $\SpopMean$ as a function of the mean $\SphyMean$ for a subset of SNPs, for which $\SphyMean$ is close to $0$.
    Indeed, polyDFE is a nearly neutral model and is not able to reliably compute selection coefficients for only strongly deleterious or only strongly advantageous SNPs.
    Sliding window of groups of $5.000$ SNPs are taken with a mean $\SphyMean$ between $-0.5$ and $0.5$ (fig.~\ref{fig:selcoeff-pop-phylo}B).
    $\SpopMean$ is estimated for each group, with some groups that might have part of same SNPs that overlap.
    The relationship between $\SpopMean$ and $\SphyMean$ is described as a second-order polynomial, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\SpopMean/\SphyMean$ for each population.
    We observe that $\SpopMean/\SphyMean$, representing the efficacy of selection at the short versus long-term is positively correlated with synonymous diversity (fig.~\ref{fig:selcoeff-pop-phylo}C), a proxy of effective population size.
    Altogether, populations with large effective population size are under more efficient selection, confirming the prediction of the nearly-neutral theory of evolution.

    As a consequence of increased selection efficacy, mutations in populations with high diversity are mechanically more likely to be deleterious (fig.~S4), especially for mutations predicted already to be deleterious at the phylogenetic scale (fig.~S5-6).
    However, and surprisingly, a decrease in the proportion of neutral mutation did not translate into a mechanical increase in the proportion of advantageous mutations.
    Instead, populations with higher diversity have a lower proportion of advantageous mutations (fig.~S7), this effect is also more pronounced for mutations predicted to be advantageous at the phylogenetic scale (red row in fig.~\ref{fig:all-pop-results}C and fig.~S8).
    This discrepancy can be explained by the fact that populations with larger diversity have already depleted the putatively advantageous mutations, and that only deleterious are remaining possible.
    In other words, being on the top of a fixed fitness landscape leaves you with only paths going downhill.
    Altogether, populations with increased diversity are showing signature of increased selection efficiency and are also less likely to generate positively selected mutations.

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure2}
        \caption{
            Panel A: Histogram of predicted selection coefficient ($\Sphy$) for each observed mutation in a sample of 8 individuals (out of 512 in the original dataset) with African descent.
            Mutations are divided in 5 classes of selection: severely deleterious (blue), deleterious (green), weakly deleterious (light green), weakly advantageous (yellow) and advantageous (red).
            Panel B: Site-frequency spectrum (SFS) represents the proportion of mutations (y-axis) with a given number of derived alleles in the population (x-axis).
            SFS are drawn for a random sample of 16 alleles (mean in solid line and standard deviation in filled color) for each class of selection and for synonymous mutations supposedly neutral (black).
            Supposedly advantageous are over-represented at higher frequency, while severally deleterious are underrepresented at high frequency, a signature of selection.
            Panel C. Histogram of predicted selection coefficient ($\Sphy$) for each possible mutation away from the ancestral human genome, weighted by the mutation rate.
            For each class of selection, the sum of histogram in this class thus gives the expected total mutation rate toward this class.
            If they are less mutations observed than expected, this class is thus undergoing purifying selection.
            Panel D. For each class of selection (and for the set of all non-synonymous mutations), information from the SFS and the expected total mutation rate are combined at the population-genetic scale to estimate the proportion of advantageous mutations $\PpolyAdv$, of nearly-neutral mutations $\PpolyNeutral$ and of deleterious mutations $\PpolyDel$.
        }
        \label{fig:homo-afr-results}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure3}
        \caption{
            Reproducibility of results shown in figure~\ref{fig:homo-afr-results} in 29 populations across 7 genera.
            Site-frequency spectra and total mutation rates are combined at the population-genetic scale to estimate the proportion of advantageous mutations ($\PpolyAdv$) in top panel, of nearly-neutral mutations ($\PpolyNeutral$) in middle panel and of deleterious mutations ($\PpolyDel$) in bottom panel.
        }
        \label{fig:all-pop-results}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure4}
        \caption{
            Panel A. For each population, proportion of advantageous ($\PpolyAdv$), nearly-neutral ($\PpolyNeutral$) and of deleterious ($\PpolyDel$) mutations estimated at the population-genetic scale.
            Estimation is performed with SNPs predicted to be weakly advantageous at the phylogenetic scale ($\divWeakAdv$).
            Panel B. Mean selection coefficient at the population scale $\SpopMean$ as a function of phylogenetic $\SphyMean$.
            Observed SNPs are sorted and groups of $5.000$ SNPs are taken as a sliding window with increasing mean $-0.5 \leq \SphyMean \leq 0.5$.
            For each group along the sliding window, $\SpopMean$ (y-axis) is estimated based on SFS and the total mutation rate, and compared to the mean $\SphyMean$ (x-axis).
            A second order polynomial is fitted for each population, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\SpopMean/\SphyMean$ given in legend.
            Panel C. Ratio $\SpopMean/\SphyMean$ as a function of the synonymous diversity for each population.
            Populations with higher diversity are showing higher ratio of ratio $\SpopMean/\SphyMean$.
        }
        \label{fig:selcoeff-pop-phylo}
    \end{figure*}

    From a structural biology perspective, we found that proteins are not optimal, with a large proportion of mutations still toward more optimal amino acids.
    Seen from another angle, it means that we could theoretically derive the fittest proteins using only the amino acid with the highest fitness.
    Another way to confirm our predictions on amino acid fitness would be to test whether fittest proteins are also thermodynamically more stable, leveraging recent developments able to compute efficiently and accurately the folding of proteins.
